---
title: "Getting and cleaning Data Swirl exercises"
author: "Hariharan"
output:
  pdf_document: default
  keep_md: yes
---
This PDF contains the code and output generated for swirl exercises in the course
Find the Getting-and-cleaning-data-swirl.Rmd in the same folder as this file to interact with the code and make changes for a better learning experience

# 1. Manipulating Data with dplyr

## Setting up the environment

```{r}
path2csv<-"C:/Users/MAHE/Documents/R/win-library/3.6/swirl/Courses/Getting_and_Cleaning_Data/Manipulating_Data_with_dplyr/2014-07-08.csv"
mydf <- read.csv(path2csv, stringsAsFactors = FALSE)
dim(mydf)
head(mydf)
# loading dplyr
library(dplyr)
packageVersion("dplyr")
```
 The first step of working with data in dplyr is to load the data into what the package authors call a 'data frame tbl' or 'tbl_df'. Use the following code to create a new tbl_df called cran:
 
 
```{r}
cran <- tbl_df(mydf)
rm("mydf")
cran
```
## using select
 
```{r}
select(cran, ip_id, package, country)
select(cran, r_arch:country)
select(cran, -time)
select(cran, -(X:size))
```
## using filter
```{r}
filter(cran, package == "swirl")
filter(cran, r_version == "3.1.1", country == "US")
filter(cran, r_version <= "3.0.2", country == "IN")
filter(cran, country == "US" | country == "IN")
filter(cran, size > 100500, r_os == "linux-gnu")
filter(cran, !is.na(r_version))
```
## using arrange

```{r}
cran2 <- select(cran, size:ip_id)
arrange(cran2, ip_id)#ip_id is in ascending order
arrange(cran2, desc(ip_id))#To do the same, but in descending order
arrange(cran2, package, ip_id)# first arrange by package names (ascending alphabetically), then by ip_id
arrange(cran2, country, desc(r_version), ip_id)
```

## using mutate

```{r}
cran3 <- select(cran, ip_id, package, size)
mutate(cran3, size_mb = size / 2^20)
mutate(cran3, size_mb = size / 2^20, size_gb = size_mb / 2^10)
mutate(cran3, correct_size = size + 1000)
```


## using summarize
```{r}
summarize(cran, avg_bytes = mean(size))

```

# 2. Grouping and Chaining with dplyr

## Setting up the environment

```{r}
path2csv<-"C:/Users/MAHE/Documents/R/win-library/3.6/swirl/Courses/Getting_and_Cleaning_Data/Manipulating_Data_with_dplyr/2014-07-08.csv"
mydf <- read.csv(path2csv, stringsAsFactors = FALSE)
dim(mydf)
head(mydf)
# loading dplyr
library(dplyr)
packageVersion("dplyr")
cran <- tbl_df(mydf)
```
## using group-by

```{r}
by_package <- group_by(cran, package)
by_package
summarize(by_package, mean(size))
```
```{r}
pack_sum <- summarize(by_package,
                      count = n(),
                      unique = n_distinct(ip_id),
                      countries = n_distinct(country),
                      avg_bytes = mean(size))
pack_sum

```

The 'count' column, created with n(), contains the total number of rows (i.e.downloads) for each package. The 'unique' column, created with n_distinct(ip_id), gives the total number of unique downloads for each
package, as measured by the number of distinct ip_id's. The 'countries'
column, created with n_distinct(country), provides the number of countries in which each package was downloaded. And finally, the 'avg_bytes' column, created with mean(size), contains the mean download size (in bytes) for each package.

 Naturally, we'd like to know which packages were most popular on the day these data were collected (July 8, 2014). Let's start by isolating the top 1% of packages, based on the total number of downloads as measured by the 'count' column.
```{r}
quantile(pack_sum$count, probs = 0.99)
top_counts <- filter(pack_sum, count > 679)
top_counts
View(top_counts)
top_counts_sorted <- arrange(top_counts, desc(count))
top_counts_sorted
```
```{r}
quantile(pack_sum$unique, probs = 0.99)
top_unique <- filter(pack_sum, unique > 465)
top_unique_sorted <- arrange(top_unique, desc(unique))
top_unique_sorted
```
## Chaining
```{r}
by_package <- group_by(cran, package)
pack_sum <- summarize(by_package,
                      count = n(),
                      unique = n_distinct(ip_id),
                      countries = n_distinct(country),
                      avg_bytes = mean(size))
top_countries <- filter(pack_sum, countries > 60)
result1 <- arrange(top_countries, desc(countries), avg_bytes)
print(result1)
```
## Same operations as above but using function call embedding
```{r}
result2 <-
  arrange(
    filter(
      summarize(
        group_by(cran,
                 package
        ),
        count = n(),
        unique = n_distinct(ip_id),
        countries = n_distinct(country),
        avg_bytes = mean(size)
      ),
      countries > 60
    ),
    desc(countries),
    avg_bytes
  )

print(result2)
```
In this script, we've used a special chaining operator, %>%
```{r}
# you read it, you can pronounce the %>% operator as
# the word 'then'.

result3 <-
  cran %>%
  group_by(package) %>%
  summarize(count = n(),
            unique = n_distinct(ip_id),
            countries = n_distinct(country),
            avg_bytes = mean(size)
  ) %>%
  filter(countries > 60) %>%
  arrange(desc(countries), avg_bytes)

# Print result to console
print(result3)
```
```{r}
# select() the following columns from cran. Keep in mind
# that when you're using the chaining operator, you don't
# need to specify the name of the data tbl in your call to
# select().
#
# 1. ip_id
# 2. country
# 3. package
# 4. size
#
# The call to print() at the end of the chain is optional,
# but necessary if you want your results printed to the
# console. Note that since there are no additional arguments
# to print(), you can leave off the parentheses after
# the function name. This is a convenient feature of the %>%
# operator.

cran %>%
  select(ip_id, country, package, size) %>%
	print


# Use mutate() to add a column called size_mb that contains
# the size of each download in megabytes (i.e. size / 2^20).
#
# If you want your results printed to the console, add
# print to the end of your chain.

cran %>%
  select(ip_id, country, package, size) %>%
  mutate(size_mb = size / 2^20)

# Use filter() to select all rows for which size_mb is
# less than or equal to (<=) 0.5.
#
# If you want your results printed to the console, add
# print to the end of your chain.

cran %>%
  select(ip_id, country, package, size) %>%
  mutate(size_mb = size / 2^20) %>%
  filter(size_mb <= 0.5)


# arrange() the result by size_mb, in descending order.
#
# If you want your results printed to the console, add
# print to the end of your chain.

cran %>%
  select(ip_id, country, package, size) %>%
  mutate(size_mb = size / 2^20) %>%
  filter(size_mb <= 0.5) %>%
  arrange(desc(size_mb))

```


# 3.Tidying Data with tidyr

## Setting up 
```{r}
library(tidyr)
# recreating dataset usend in lesson
students<- data.frame("grade" =c('A','B','C','D','E') , "male" = as.integer(c(5,4,8,4,5)), "female" = as.integer(c(3,1,6,5,5)))

students

```
```{r}
gather(students, sex, count, -grade)
```

```{r}
students2<- data.frame("grade" =c('A','B','C','D','E') , "male_1" = as.integer(c(7,4,7,8,8)), "female_1" = as.integer(c(0,0,4,2,4)),"male_2" = as.integer(c(5,5,5,8,1)), "female_2" = as.integer(c(8,8,6,1,0)))

students2
```
```{r}
res <- gather(students2, sex_class, count, -grade)
res
separate(res, sex_class, c("sex", "class"))
```
## using chaining
```{r}
students2 %>%
  gather(sex_class, count, -grade) %>%
  separate(sex_class, c("sex", "class")) %>%
  print
```

```{r}
students3<- data.frame(
  "name" = c("Sally","Sally","Jeff","Jeff","Roger","Roger","Karen","Karen","Brian","Brian"),
  "test" = c("midterm","final","midterm","final","midterm","final","midterm","final","midterm","final"),
  "class1" = c("A","C",NA,NA,NA,NA,NA,NA,"B","B"),
  "class2" = c(NA,NA,"D","E","C","A",NA,NA,NA,NA),
  "class3" = c("B","C",NA,NA,NA,NA,"C","C",NA,NA),
  "class5" = c(NA,NA,NA,NA,"B","A",NA,NA,"A","C"),
  stringsAsFactors = FALSE
)
students3
```
```{r}
# Call gather() to gather the columns class1
# through class5 into a new variable called class.
# The 'key' should be class, and the 'value'
# should be grade.
#
# tidyr makes it easy to reference multiple adjacent
# columns with class1:class5, just like with sequences
# of numbers.
#
# Since each student is only enrolled in two of
# the five possible classes, there are lots of missing
# values (i.e. NAs). Use the argument na.rm = TRUE
# to omit these values from the final result.
#
# Remember that when you're using the %>% operator,
# the value to the left of it gets inserted as the
# first argument to the function on the right.
#
# Consult ?gather and/or ?chain if you get stuck.
#
students3 %>%
  gather(class, grade, class1:class5, na.rm = TRUE) %>%
  print

```
```{r}
# This script builds on the previous one by appending
# a call to spread(), which will allow us to turn the
# values of the test column, midterm and final, into
# column headers (i.e. variables).
#
# You only need to specify two arguments to spread().
# Can you figure out what they are? (Hint: You don't
# have to specify the data argument since we're using
# the %>% operator.
#
students3 %>%
  gather(class, grade, class1:class5, na.rm = TRUE) %>%
  spread(test, grade) %>%
  print
```
```{r}
library(readr)
parse_number("class5")
```

```{r}
# We want the values in the class columns to be
# 1, 2, ..., 5 and not class1, class2, ..., class5.
#
# Use the mutate() function from dplyr along with
# parse_number(). Hint: You can "overwrite" a column
# with mutate() by assigning a new value to the existing
# column instead of creating a new column.
#
# Check out ?mutate and/or ?parse_number if you need
# a refresher.
#
students3 %>%
  gather(class, grade, class1:class5, na.rm = TRUE) %>%
  spread(test, grade) %>%
  mutate(class = parse_number(class)) %>%
  print
```

```{r}
students4 <- data.frame(
  "id" = as.integer(c(168,168,588,588,710,710,731,731,908,908)),
  "name" = c("Brian","Brian","Sally","Sally","Jeff","Jeff","Roger","Roger","Karen","Karen"),
  "sex" = c("F","F","M","M","M","M","F","F","M","M"),
  "class" = as.integer(c(1,5,1,3,2,4,2,5,3,4)),
  "midterm" = c("B","A","A","B","D","A","C","B","C","A"),
  "final" = c("B","C","C","C","E","C","A","A","C","A"),
  stringsAsFactors = FALSE
  
)
students4
```
```{r}
# selecting the id, name, and sex column from students4
# and storing the result in student_info.
#
student_info <- students4 %>%
  select(id, name, sex) %>%
  print
# Add a call to unique() below, which will remove
# duplicate rows from student_info.
#
# Like with the call to the print() function below,
# you can omit the parentheses after the function name.
# This is a nice feature of %>% that applies when
# there are no additional arguments to specify.
#
student_info <- students4 %>%
  select(id, name, sex) %>%
  unique %>%
  print

# select() the id, class, midterm, and final columns
# (in that order) and store the result in gradebook.
#
gradebook <- students4 %>%
  select(id, class, midterm, final) %>%
  print
```

```{r}
passed<- data.frame(
  "name" = c("Brian","Roger","Roger","Karen"),
  "class" = as.integer(c(1,2,5,4)),
  "final" = c("B","A","A","A"),
  stringsAsFactors = FALSE
  )
failed <- data.frame(
  "name" = c("Brian","Sally","Sally","Jeff","Jeff","Karen"),
  "class" = as.integer(c(5,1,3,2,4,3)),
  "final" = c("C","C","C","E","C","C"),
   stringsAsFactors = FALSE
)
passed
failed
```
```{r}
passed <- passed %>% mutate(status = "passed")
failed <- failed %>% mutate(status = "failed")
bind_rows(passed, failed)
 
```

# 4. Dates and Times with lubridate

```{r}
library(lubridate)
this_day <- today()
this_day
month(this_day)
wday(this_day)
wday(this_day, label = TRUE)
this_moment <- now()
this_moment
second(this_moment)
my_date <- ymd("1989-05-17")
my_date
class(my_date)
ymd("1989 May 17")
mdy("March 12, 1975")
dmy(25081985)
ymd("1920/1/2")
dt1 <- "2014-08-23 17:23:02"
ymd_hms(dt1)
hms("03:22:14")
dt2<-c("2014-05-14","2014-09-22","2014-07-11")
ymd(dt2)
update(this_moment, hours = 8, minutes = 34, seconds = 55)
this_moment
this_moment <- update(this_moment, hours = 10, minutes = 16, seconds = 0)
nyc <- now("America/New_York")
depart <- nyc + days(2)
depart <- update(depart, hours = 17, minutes = 34)
arrive <- depart + hours(15) + minutes(50)
arrive <- with_tz(arrive, "Asia/Hong_Kong")
last_time <- mdy("June 17, 2008", tz = "Singapore")
how_long <- interval(last_time, arrive)
as.period(how_long)
```
