---
title: "R Programming Swirl Exercises"
author: "Hariharan"
date: "5/11/2020"
output:
  pdf_document: default
  keep_md: yes
---
This PDF contains the code and output generated for swirl exercises in the course
Find the R programming.Rmd in the same folder as this file to interact with the code and make changes for a better learninng experience


# 1. Basic Building Blocks

## Assignment operator
```{r}
x<-5+7
x
y<-x-3
y
```

## Concatenate/combine function
```{r}
z<-c(1.1 , 9,3.14)
z
c(z,555,z)

```

## Using numeric vectors in arithmetic expressions
```{r}
z*2+100
#using sqrt function
my_sqrt<-sqrt(z-1)
my_sqrt

my_div<-z/my_sqrt
my_div

```

## Vector Recycling
```{r}
c(1,2,3,4)+c(0,10)
c(1,2,3,4) + c(0,10,100)

```
# 2. Work space and files
N/A


# 3. Sequence of Numbers

## using :
```{r}
pi:10

15:1

```

## using seq()
```{r}
seq(1,20)
seq(0,10,by=0.5)
my_seq<-seq(5,10,length=30)
length(my_seq)
seq(along.with = my_seq)
seq_along(my_seq)

```
## Repeating Sequences
```{r}
rep(0, times = 40)
rep(c(0,1,2),times=10)
rep(c(0,1,2),each=10)

```

# 4. Vectors

## Basic operations
```{r}
num_vect<-c(0.5,55,-10,6)
tf<-num_vect <1
tf
num_vect>=6

```
## Character vectors

```{r}
my_char <- c("My", "name", "is")
my_char
#to join the vector
paste(my_char, collapse = " ")
#concatenation
my_name <- c(my_char, "Swirl")
my_name
paste(my_name, collapse = " ")
paste("Hello", "world!", sep = " ")
paste(1:3, c("X", "Y", "Z"), sep = "")
paste(LETTERS, 1:4, sep = "-")

```

# 5. Missing Values

## finding NA
```{r}
x<-c(44,NA,5,NA)
x*3
# creating a dataset with random NA's
y<-rnorm(1000)
z<-rep(NA,1000)
my_data<-sample(c(y,z),100)
sum(my_data)
#my_na contains bool values representing if that element is na or not
my_na<-is.na(my_data)
 

```
## finding NaN
```{r}
# instances where Nan is generated
0/0
Inf-Inf
```

# 6. Subsetting Vectors
## forming random data set
```{r}
x<-sample(c(rnorm(20),rep(NA,20)),40)
x[1:10]
```
## handling NA
```{r}
# extracting all NA
x[is.na(x)]
#filtering missing values
y<-x[!is.na(x)]
y
```

## more subsetting
```{r}
y[y>0]
x[!is.na(x) & x > 0]
x[c(3, 5, 7)]
x[c(-2, -10)]
 x[-c(2, 10)]
```
## vectors with named elements

```{r}
vect <- c(foo = 11, bar = 2, norf = NA)
vect
names(vect)
#adding names
vect2 <- c(11, 2, NA)
names(vect2) <- c("foo", "bar", "norf")
identical(vect, vect2)
#filtering
vect["bar"]
vect[c("foo", "bar")]

```

# 7. Matrices and Data Frames

```{r}
my_vector <- 1:20
dim(my_vector)
dim(my_vector) <- c(4, 5)
dim(my_vector) <- c(4, 5)
 my_vector
 my_matrix <- my_vector
 

```

## Using the matrix function
```{r}
my_matrix2 <- matrix(1:20, nrow=4, ncol=5)
```

## Labeling 
```{r}
patients <- c("Bill", "Gina", "Kelly", "Sean")
cbind(patients, my_matrix)
```

## Constructing a data frame
```{r}
my_data <- data.frame(patients, my_matrix)
my_data
cnames <- c("patient", "age", "weight", "bp", "rating", "test")
colnames(my_data) <- cnames
my_data
```

# 8. Logic

## Some basic operations
```{r}
TRUE == TRUE
 7 == 9
 7 > 9
 7 < 9
 7>=9
 7!=9
 
```
You can use the `&` operator to evaluate AND across a vector. The `&&`
version of AND only evaluates the first member of a vector. 

```{r}
TRUE && c(TRUE, FALSE, FALSE)
TRUE & c(TRUE, FALSE, FALSE)
TRUE | c(TRUE, FALSE, FALSE)
TRUE || c(TRUE, FALSE, FALSE)
```

## Logical functions in R
```{r}
isTRUE(6 > 4)
xor(5 == 6, !FALSE)
ints <- sample(10)
ints > 5
which(ints > 7)
any(ints < 0)
```

# 9. Functions

N/A

# 10. lapply() and sapply()

## code to download the data set for this section
```{r}
y<- "https://archive.ics.uci.edu/ml/machine-learning-databases/flags/flag.data"

flags <- read.table(y, header = FALSE, sep = ",")
names(flags) <- c("country", "landmass", "zone", "area", "population", "language", 
              "religion", "bars", "stripes", "colors", "red", "green", "blue",
              "gold", "white", "black", "orange","mainhue", "circles", "crosses", 
              "saltires", "quarters", "sunstars", "crescents", "triangle", "icon",
              "animate", "text", "topleft", "botright")


```


```{r}
head(flags)
dim(flags)
class(flags)
```

The lapply() function takes a list as input, applies a function to each
element of the list, then returns a list of the same length as the original
one. Since a data frame is really just a list of vectors (you can see this
with as.list(flags)), we can use lapply() to apply the class() function to
each column of the flags dataset. 

```{r}
cls_list <- lapply(flags, class)
cls_list
as.character(cls_list)
```
if we use saaply the result is simplified in this case 
list will be character vector automatically


```{r}
cls_vect <- sapply(flags, class)
class(cls_vect)
```
 To see how many flags have orange in them we use
```{r}
sum(flags$orange)
```
Now we want to repeat this operation for each of the colors recorded in the dataset.

First, use flag_colors <- flags[, 11:17] to extract the columns containing
the color data and store them in a new data frame called flag_colors. (Note the comma before 11:17. This subsetting command tells R that we want all rows, but only columns 11 through 17.)


```{r}
flag_colors <- flags[, 11:17]
head(flag_colors)
lapply(flag_colors, sum)
sapply(flag_colors, sum)

```

To find mean use of each color

```{r}
sapply(flag_colors, mean)
```

```{r}
flag_shapes <- flags[, 19:23]

lapply(flag_shapes, range)
shape_mat <- sapply(flag_shapes, range)
 shape_mat
```
using unique() 

```{r}
unique(c(3, 4, 5, 5, 5, 6, 6))
unique_vals <- lapply(flags, unique)
unique_vals
```



```{r}
sapply(unique_vals, length)
```

# 11. vapply and tapply
in vapply we can specify the output type instead of letting R guess the format
```{r}
vapply(flags, class, character(1))
```
 using tapply()
 
```{r}
table(flags$landmass)
table(flags$animate)
```
If you take the arithmetic mean of a bunch of 0s and 1s, you get the
proportion of 1s. Use tapply(flags$animate, flags$landmass, mean) to apply the mean function to the 'animate' variable separately for each of the six landmass groups, thus giving us the proportion of flags containing an animate image WITHIN each landmass group.



```{r}
tapply(flags$animate, flags$landmass, mean)
```
Similarly, we can look at a summary of population values (in round millions)
for countries with and without the color red on their flag with
tapply(flags$population, flags$red, summary).

```{r}
tapply(flags$population, flags$red, summary)
```


# 12. Looking at data

loading a sample dataset for this lesson
```{r}
data("mtcars")

```

Some functions to look at data

```{r}
dim(mtcars)
nrow(mtcars)
ncol(mtcars)
object.size(mtcars)
names(mtcars)
head(mtcars)
tail(mtcars,15)
summary(mtcars)
str(mtcars)
```


# 13. Simulation
Let's simulate rolling four six-sided dice:
```{r}
sample(1:6, 4, replace = TRUE)
```



```{r}
sample(1:20, 10)
```

```{r}
LETTERS
```

Let the value 0 represent tails and the value 1 represent heads. Use sample() to draw a sample of size 100 from the vector c(0,1), with replacement. Since the coin is unfair, we must attach specific probabilities to the values 0 (tails) and 1 (heads) with a fourth argument, prob = c(0.3, 0.7). Assign the result to a new variable called flips.

```{r}
flips <- sample(c(0,1), 100, replace = TRUE, prob = c(0.3, 0.7))
flips
sum(flips)
```

 A binomial random variable represents the number of 'successes' (heads) in a given number of independent 'trials' (coin flips). Therefore, we can generate a single random variable that represents the number of heads in 100 flips of our unfair coin using rbinom(1, size = 100, prob = 0.7). Note that you only specify the probability of 'success' (heads) and NOT the probability of 'failure' (tails). 

```{r}
rbinom(1, size = 100, prob = 0.7)
flips2 <- rbinom(100, size = 1, prob = 0.7)
flips2
sum(flips2)
```

The standard normal distribution has mean 0 and standard deviation 1. As you can see under the 'Usage' section in the documentation, the default values for the 'mean' and 'sd' arguments to rnorm() are 0 and 1, respectively. Thus,rnorm(10) will generate 10 random numbers from a standard normal distribution.


```{r}
rnorm(10)
# with a mean of 100 and a standard deviation of 25
rnorm(10, 100, 25)
```
what if we want to simulate 100 *groups* of random numbers, each containing 5 values generated from a Poisson distribution with mean 10? Let's start with one group of 5 numbers, then I'll show you how to repeat the operation 100 times in a convenient and compact way.

```{r}
rpois(5, 10)
my_pois <- replicate(100, rpois(5, 10))
my_pois
cm <- colMeans(my_pois)
hist(cm)
```

# 14. Dates and Times
## Dates
```{r}
d1 <- Sys.Date()
unclass(d1)
d1
d2 <- as.Date("1969-01-01")
unclass(d2)

```


## Times

```{r}
t1 <- Sys.time()
t1
class(t1)
unclass(t1)
t2 <- as.POSIXlt(Sys.time())
t2
unclass(t2)
str(unclass(t2))
t2$min
```

## other functions to work with date and time

```{r}
weekdays(d1)
months(t1)
quarters(t2)
```

## Processing from string

```{r}
t3 <- "October 17, 1986 08:24"
t4 <- strptime(t3, "%B %d, %Y %H:%M")
t4
```


```{r}
Sys.time() > t1
Sys.time() - t1
difftime(Sys.time(), t1, units = 'days')
```




# 15. Base Graphics

```{r}
data(cars)
head(cars)
plot(cars)
plot(x = cars$speed, y = cars$dist,xlab= "Speed" , ylab = "Stopping distance", main = "My Plot", sub = "My Plot Subtitle" , col = 2 , pch = 2 )

```

```{r}
data(mtcars)
boxplot(formula = mpg ~ cyl, data = mtcars)
hist(mtcars$mpg)

```

