---
title: "Programming Assignment Air quality"
author: "Hariharan"
output:
  pdf_document: default
  keep_md: yes
---
This PDF contains the code and output generated for programming assignment 3 in the course
Find the week_2_prog_assignment.Rmd in the same folder as this file to interact with the code and make changes for a better learning experience
Download the file ProgAssignment3-data.zip file containing the data for Programming Assignment 3 from
the Coursera web site. Unzip the file in a directory that will serve as your working directory. When you
start up R make sure to change your working directory to the directory where you unzipped the data.

# 1 Plot the 30-day mortality rates for heart attack
```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11])
```

# 2 Finding the best hospital in a state
Write a function called best that take two arguments: the 2-character abbreviated name of a state and an
outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector
with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome
in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can
be one of “heart attack”, “heart failure”, or “pneumonia”. Hospitals that do not have data on a particular
outcome should be excluded from the set of hospitals when deciding the rankings.

```{r}
best <- function(state, outcome) {
  ## Read outcome data
  ## Check that state and outcome are valid
  ## Return hospital name in that state with lowest 30-day death
  ## rate
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  fd   <- as.data.frame(cbind(data[, 2],   # hospital
                              data[, 7],   # state
                              data[, 11],  # heart attack
                              data[, 17],  # heart failure
                              data[, 23]), # pneumonia
                        stringsAsFactors = FALSE)
  colnames(fd) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
  
  if(!state %in% fd[, "state"])
  {
      stop('invalid state')
  }
  else if(!outcome %in% c("heart attack", "heart failure", "pneumonia"))
  {
    stop('invalid outcome')
  }
  else 
  {
    si <- which(fd[, "state"] == state)
    ts <- fd[si, ]    # extracting data for the called state
    oi <- as.numeric(ts[,outcome])
    min_val <- min(oi, na.rm = TRUE)
    result  <- ts[, "hospital"][which(oi == min_val)]
    output  <- result[order(result)]
  }
  return(output)
  
}

```
## Sample usage

```{r}
best("SC", "heart attack")  
```


# 3 Ranking hospitals by outcome in a state
Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a
state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num).
The function reads the outcome-of-care-measures.csv file and returns a character vector with the name
of the hospital that has the ranking specified by the num argument. For example, the call
rankhospital("MD", "heart failure", 5)
would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate
for heart failure. The num argument can take values “best”, “worst”, or an integer indicating the ranking
(smaller numbers are better). If the number given by num is larger than the number of hospitals in that
state, then the function should return NA. Hospitals that do not have data on a particular outcome should
be excluded from the set of hospitals when deciding the rankings.

```{r}
rankhospital <- function(state, outcome, rank = "best"){
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  fd   <- as.data.frame(cbind(data[, 2],  # hospital
                              data[, 7],  # state
                              data[, 11],  # heart attack
                              data[, 17],  # heart failure
                              data[, 23]), # pneumonia
                        stringsAsFactors = FALSE)
  colnames(fd) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
  
  ## Check that state and outcome are valid
  if (!state %in% fd[, "state"]) {
    stop('invalid state')
  } else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")){
    stop('invalid outcome')
  } else if (is.numeric(rank)) {
    si <- which(fd[, "state"] == state)
    ts <- fd[si, ]                     # extracting dataframe for the called state
    ts[, eval(outcome)] <- as.numeric(ts[, eval(outcome)])
    ts <- ts[order(ts[, eval(outcome)], ts[, "hospital"]), ]
    output <- ts[, "hospital"][rank]
  } else if (!is.numeric(rank)){
    if (rank == "best") {
      output <- best(state, outcome)
    } else if (rank == "worst") {
      si <- which(fd[, "state"] == state)
      ts <- fd[si, ]    
      ts[, eval(outcome)] <- as.numeric(ts[, eval(outcome)])
      ts <- ts[order(ts[, eval(outcome)], ts[, "hospital"], decreasing = TRUE), ]
      output <- ts[, "hospital"][1]
    } else {
      stop('invalid rank')
    }
  }
  return(output)
}

```
## Sample Usage
```{r}
print(rankhospital("NC", "heart attack", "worst"))
```

# 4 Ranking hospitals in all states
Write a function called rankall that takes two arguments: an outcome name (outcome) and a hospital ranking (num). The function reads the outcome-of-care-measures.csv file and returns a 2-column data frame
containing the hospital in each state that has the ranking specified in num. For example the function call
rankall("heart attack", "best") would return a data frame containing the names of the hospitals that
are the best in their respective states for 30-day heart attack death rates. The function should return a value
for every state (some may be NA). The first column in the data frame is named hospital, which contains
the hospital name, and the second column is named state, which contains the 2-character abbreviation for
the state name. Hospitals that do not have data on a particular outcome should be excluded from the set of
hospitals when deciding the rankings.

```{r}
rankall <- function(outcome, num = "best"){
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  fd   <- as.data.frame(cbind(data[, 2],  # hospital
                              data[, 7],  # state
                              data[, 11],  # heart attack
                              data[, 17],  # heart failure
                              data[, 23]), # pneumonia
                        stringsAsFactors = FALSE)
  colnames(fd) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
  fd[, eval(outcome)] <- as.numeric(fd[, eval(outcome)])
  
  ## Check that state and outcome are valid
  
  if (!outcome %in% c("heart attack", "heart failure", "pneumonia")){
    stop('invalid outcome')
  } else if (is.numeric(num)) {
    by_state <- with(fd, split(fd, state))
    ordered  <- list()
    for (i in seq_along(by_state)){
      by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                           by_state[[i]][, "hospital"]), ]
      ordered[[i]]  <- c(by_state[[i]][num, "hospital"], by_state[[i]][, "state"][1])
    }
    result <- do.call(rbind, ordered)
    output <- as.data.frame(result, row.names = result[, 2], stringsAsFactors = FALSE)
    names(output) <- c("hospital", "state")
  } else if (!is.numeric(num)) {
    if (num == "best") {
      by_state <- with(fd, split(fd, state))
      ordered  <- list()
      for (i in seq_along(by_state)){
        by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                             by_state[[i]][, "hospital"]), ]
        ordered[[i]]  <- c(by_state[[i]][1, c("hospital", "state")])
      }
      result <- do.call(rbind, ordered)
      output <- as.data.frame(result, stringsAsFactors = FALSE)
      rownames(output) <- output[, 2]
    } else if (num == "worst") {
      by_state <- with(fd, split(fd, state))
      ordered  <- list()
      for (i in seq_along(by_state)){
        by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                             by_state[[i]][, "hospital"], 
                                             decreasing = TRUE), ]
        ordered[[i]]  <- c(by_state[[i]][1, c("hospital", "state")])
      }
      result <- do.call(rbind, ordered)
      output <- as.data.frame(result, stringsAsFactors = FALSE)
      rownames(output) <- output[, 2]
    } else {
      stop('invalid num')
    }
  }
  return(output)
}
```
## Sample Useage
```{r}
print(rankall("heart attack", "worst"))
```

# Quiz


```{r}

best("SC", "heart attack")

best("NY", "pneumonia")

best("AK", "pneumonia")

rankhospital("NC", "heart attack", "worst")

rankhospital("WA", "heart attack", 7)

rankhospital("TX", "pneumonia", 10)

rankhospital("NY", "heart attack", 7)

r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)

r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)

r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)


```


